### 2021-07-01 Documents/TODO.Rmd
bayes factor first trial

---
title: 'TODO'
output:
  bookdown::html_document2: 
    toc: true
    toc_depth: 3
    toc_float: true
    collapsed: false
    # css: ["css/custom.css"]
linestretch: 1.5
number_sections: true
# bibliography: "../Biblio-attachment.bib"
# csl: apa.csl
link-citations: true 
editor_options: 
  chunk_output_type: console
---

# Bayes Factor Encompassing Priors

- [ ] Reparametrize model according to: 
\begin{align}
&\beta_{eq} = R_{eq}\theta - r_{eq},\\
&\beta_{ineq} = R_{ineq}\theta - r_{ineq}.
\end{align}
Doing this we get informative hypotheses as $H_i: \beta_{eq} = 0,\ \beta_{ineeq}>0$, so prior specification is semplified and with simmetric prior we do  not favoir any constraint. Exception for range constraints of type $a<\theta<b$.

- [X] Marginal densities of a multivariate normal distribution are obtained simply dropping irreleevantt variables ([link math-stackexchange](https://math.stackexchange.com/questions/2528201/find-the-marginal-distributions-pdfs-of-a-multivariate-normal-distribution))