### 2021-09-14 Documents/Bookdown/index.Rmd
add template github pag bookdown

--- 
title: "Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach"
subtitle: "Supplemental Material"
author: "Claudio Zandonella Callegher, Tatiana Marci, Pietro De Carli, and Gianmarco Altoè"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: ["../Biblio-attachment.bib"]
csl: "../apa.csl"
link-citations: yes
github-repo: ClaudioZandonella/Attachment
description: "This the Suppleemental Material of the article ..."
cover-image: "images/logo_psicostat.png"
header-includes: 
  - \usepackage{titlepic}
  - \titlepic{\includegraphics[width=\textwidth]{images/logo_psicostat.pdf}}
url: 'https\://ClaudioZandonella.github.io/Attachment/'
params:
  format: html
---

```{r, echo = FALSE}
library(kableExtra)
library(ggplot2)

# Chunks settings
knitr::opts_chunk$set(echo = FALSE,
                      # Plot settings
                      # dev = "tikz", dev.args=list(pointsize=12),fig.align='center',
                      # fig.height=3, fig.width=5, fig.pos = "!ht",

                      # Code output width
                      tidy=TRUE, tidy.opts = list(width.cutoff = 80),
                      # comment = NA, prompt = TRUE

                      # Cache options
                      cache = FALSE, autodep=TRUE)

# Console output width
options(width = 80)

# Chunk theme
thm = knitr::knit_theme$get("bclear")
knitr::knit_theme$set(thm)
knitr::opts_chunk$set(background = c(.98, .98, 1))

# Option KableExtra
# options(knitr.kable.NA = '')

## ggplot settings
theme_set(theme_classic()+
            theme(text = element_text(size=12)))

devtools::load_all("../../")
source("../Utils_report.R")

#----    load drake    ----

drake_load_paper()
```

# Prerequisites {-}

trial