### 2021-09-17 Documents/Bookdown/02-ext.Rmd

# Externalizing Problems {#ext}

blah blaah

[@guApproximatedAdjustedFractional2018]