### 2021-09-21 Documents/Bookdown/index.Rmd

--- 
title: "Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach"
subtitle: "Supplemental Material"
author: "[Claudio Zandonella Callegher](https://claudiozandonella.netlify.app/), Tatiana Marci, Pietro De Carli, and Gianmarco Altoè"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: ["../Biblio-attachment.bib"]
csl: "../apa.csl"
link-citations: yes
github-repo: ClaudioZandonella/Attachment
description: "This the Supplemental Material of the article “Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach”"
cover-image: "images/logo_psicostat.png"
header-includes: 
  - \usepackage{titlepic}
  - \titlepic{\includegraphics[width=\textwidth]{images/logo_psicostat.pdf}}
url: 'https\://ClaudioZandonella.github.io/Attachment/'
params:
  format: html
---

```{r, echo = FALSE}
library(kableExtra)
library(ggplot2)

# Chunks settings
knitr::opts_chunk$set(echo = FALSE,
                      # Plot settings
                      fig.align='center',
                      # dev = "tikz", dev.args=list(pointsize=12),
                      # fig.height=3, fig.width=5, fig.pos = "!ht",

                      # Code output width
                      tidy=TRUE, tidy.opts = list(width.cutoff = 80),
                      # comment = NA, prompt = TRUE,
                      collapse = TRUE,

                      # Cache options
                      cache = FALSE, autodep=TRUE)

# Console output width
options(width = 80)

# Chunk theme
thm = knitr::knit_theme$get("bclear")
knitr::knit_theme$set(thm)
knitr::opts_chunk$set(background = c(.98, .98, 1))

# Option KableExtra
# options(knitr.kable.NA = '')

## ggplot settings
theme_set(theme_bw()+
            theme(text = element_text(size=12)))

devtools::load_all("../../")
source("../Utils_report.R")

#----    load drake    ----

drake_load_bookdown()
```

# Introduction {-}

This is the Supplemental Material of the article “*Evaluating Informative Hypotheses With Equality and Inequality Constraint: The Bayes Factor Encompassing Prior Approach*” [TODO: add cit]. The aim of the paper is to provide a clear and detailed description of the Bayes Factor with the encompassing prior approach considering an applied example.

## Study Summary {-}

When conducting a study, researchers usually have expectations based on hypotheses or theoretical perspectives they want to evaluate. Equality and inequality constraints on the model parameters are used to formalize researchers' expectations or theoretical perspectives into the so-called “*informative hypotheses*”. However, traditional statistical approaches, such as the Null Hypothesis Significance Testing (NHST) or the model comparison using information criteria (e.g., AIC and BIC), are unsuitable for testing complex informative hypotheses. An alternative approach is to use the Bayes Factor. In particular, the Bayes Factor based on the “*encompassing prior approach*” allows researchers to easily evaluate complex informative hypotheses. In this paper we provide a clear and detailed introduction to this approach. We describe all the steps and elements involved in the formalization of informative hypotheses and computation of the Bayes Factor following the encompassing prior approach. Subsequently, we provide an application of this method to a real case scenario comparing the different theoretical perspectives regarding the role of mother attachment and father attachment on children's socio-emotional development.

## Supplemental Material Structure {-}

In the supplemental material we provide the complete statistical analyses of the case study regarding attachment. For more information regarding the Bayes Factor with encompassing prior approach or the formalization of theoretical perspectives regarding the role of mother attachment and father attachment on children socio-emotional development, see the main article [TODO: add link article]. The supplemental material is structure in three parts:

#### 1) Presentation {-}

In this section, we describe the aim of the study and the sample included in the study.

- **Chapter \@ref(case-study) - Case Study.** We describe the sample included in the study and present the results of the cluster analysis to classify the different attachment patterns.

#### 2) Exteranlizing Problems {-}

In this section, the analyses to evaluate the different role of mother attachment and father attachment on children's externalizing problems are presented following three different approaches.

- **Chapter \@ref(model-choice-ext) - Models Family Choice.**  We discuss the appropriate models family to to take into account data characteristics.
- **Chapter \@ref(nhst-ext) - NHST.**  Analysis are conducted following the traditional Null Hypothesis Significance Testing (NHST).
- **Chapter \@ref(model-comparison-ext) - Model Comparison.**  Analysis are conducted according to the  Model Comparison approach using the AIC and BIC criteria.
- **Chapter \@ref(BF-ext) - Bayes Factor.**  Analysis are conducted following the Bayes Factor with encompassing prior approach.
- **Chapter \@ref(conclusion-ext) - Conclusion.**  Results of the three approaches are discussed.

#### 3) Interanlizing Problems {-}

- **Chapter \@ref(int) - Internalizing Problems.** The analyses regarding internalizing problems are presented. The analyses follow the same structure as the case of externalizing problems. Thus, only differences will be highlighted.

## GitHub {-}

All the materials are available at the GitHub repository https://github.com/ClaudioZandonella/Attachment.

Refer to the `README` file for further information.