### 2021-11-28 README.Rmd
update read me

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

# Evaluating Informative Hypotheses with Equality and Inequality Constraints:\\ A Tutorial Using the Bayes Factor via the Encompassing Prior Approach

In this repository we collect all the materials of the study “*Evaluating informative hypotheses with equality and inequality constraints: A tutorial using the Bayes factor via the encompassing prior approach*”. The paper aims to provide a clear and detailed description of the Bayes factor with the encompassing prior approach considering an applied example regarding attachment theory.

The pre-print version of the article is available at [TODO: add link]. The Supplemental Material with all the analyses is available at https://claudiozandonella.github.io/Attachment/.

## Project Structure

The project structure is based on the traditional R-package project structure, but the repository is not meant to be installed as a package (see Section [Reproducibility](#reproducibility) on how to run the analyses). R-package project structure, however, allows us to get advantage of some functionalities:

- `DESCRIPTION` file is used to track project metadata and manage project dependencies (note that we also used `renv` for package dependencies, see Section below)
- `devtools::load_all()` allows to easily load all functions defined to run the analysis (note that we followed the *functional programming paradigm*, see Section below)

Moreover, the analysis workflow is managed using `drake` (see Section below) and include other (*non-standard*) folders to organize the scripts for the analysis and for the produced outputs (i.e., paper and supplemental material).

The actual repository folder structure is:

- `Analysis/` - includes the file `Analysis_Drake.R` used to run the analysis workflow using `drake`
- `Data/` - includes the data used in the analyses
- `Documents/` - includes all the scripts and utility files for the creation of the outputs. In particular,
  - `Bookdown/` - the Supplemental Material available online
  - `Paper/` - the pre-print version of the article
- `R/` - includes all the `.R` scripts with the functions used in the analyses
- `docs/` - includes `bookdown` outputs used to build the GitHub Pages site
- `renv/` - utility folder used by `renv`

## Reproducibility {#reproducibility}

To guarantee the reproducibility of the results, the R-package`drake` is used to manage the analysis workflow and to enhance the readability and transparency of the analysis. To know more about `drake` consider the [official Git-hub page](https://github.com/ropensci/drake) or the [user manual](https://books.ropensci.org/drake/). Summarizing, using `drake` the user defines the plan of the analysis where each step in the analysis is defined through functions. Functions can be appropriately defined to obtain desired targets (i.e., R-output with results of interests) and they are declared in another script. Subsequently, `drake` manages the whole analysis recognizing the dependency structure of the different targets. When any change is made to the code `drake` evaluates the analysis and updates the results. Following the *functional programming paradigm* (i.e., defining functions for each step of the analysis) allows us to avoid "*coping and paste*" in the code, it makes debugging easier, and it facilitates the reading of the code.

Moreover, the R-package `renv` is used to manage the dependencies of the R-packages used in the analysis. The `renv` package allows the creation of an isolated, portable, and reproducible environment where the analyses are run. To know more about `renv` consider the [official documentation](https://rstudio.github.io/renv/articles/renv.html). However, `renv` is limited as it can not handle different R versions.

To overcome this issue, we also provide a `Dockerfile` to build a Docker image with R version 4.1.0. To know more about Docker see [official documentation](https://www.docker.com/) or 
[rOpenSci Labs tutorial](https://jsta.github.io/r-docker-tutorial/).

### Run the Analysis

To reproduce the analysis:

1. Download the repository on your local machine
    ```{bash, eval = FALSE}
    git clone https://github.com/ClaudioZandonella/Attachment
    ```
1. Open the R project by double-clicking the file `Attachment.Rproj` you can find in the project main directory. A new R-studio session will open.
1. Run `renv::restore()` to install the project's dependencies (have a coffee, it takes some time).
1. Now the required package will be loaded. Restart the R session (`command`/`ctrl` + `shift` + `f10`) so `devtools::load_all()` is automatically run (according to `.Rprofile`) to load all functions used in the analysis defined in the `R/` folder.

    > Note that also `source("Documents/Utils_report.R")` is automatically run to load report utility functions (functions to load analysis results and to compile the bookdown).

1. Open `Analysis/Analysis_Drake.R` and run each line (one by one) to run the analysis using drake. First, the analysis environment is set together with the analysis plan (available at `R/Plan.R`). Next, the analysis targets are computed (have a coffee, it takes a long time). Finally, the analysis results are loaded and briefly presented.

    > Note that you can access the analysis targets using the functions `drake::loadd(<name_target>)`, or load all the results with `drake_load_all()`.

    > It could happen that sometimes the session stays idle when fitting `brms` models without throwing errors. In this case, it is necessary to force restarting the R session and rerun the analysis. An error warning message will appear: `Error: drake's cache is locked`. Just follow the instructions provided by drake to force unlock of the cache  `drake::drake_cache("...")$unlock()`.

1. To compile the pre-print version of the article (possible after the analysis), open `Documents/Paper/Paper.Rnw` and compile the file using `knitr` and `pdfLaTeX` (see [project options](https://support.rstudio.com/hc/en-us/articles/200532247-Weaving-Rnw-Files-in-the-RStudio-IDE)). Note that, as usual for LaTeX files, the first run will fail. Run again to obtain the actual output. 
1. To compile the Supplemental Material, run `make_my_book()`. The resulting output is saved in the `docs/` folder.

#### Docker

To run the analysis in a Docker container, first, download the repository on your local machine. Next, build the image according to the `Dockerfile` present in the project main directory, using
```{bash, eval = FALSE}
docker build -t attachment .
```

> It is important to build the image before running the analysis

Next, you can run the container using
```{bash, eval = FALSE}
docker run --rm -p 8787:8787 -e PASSWORD="<your-PW>" attachment
```

Now, open the folder `Attachment/` and run the analysis following the instructions above from step 2.